---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(comparator)
```

# comparator: Similarity and Distance Measures in R

<!-- badges: start -->
<!-- badges: end -->

comparator implements similarity and distance measures for clustering and 
record linkage applications. It includes measures for comparing strings as 
well as numeric vectors. Where possible, measures are implemented in C/C++ 
to ensure fast performance.

## Supported measures

### String measures:

#### Edit-based:

* `Levenshtein()`: Levenshtein distance/similarity
* `DamerauLevenshtein()` Damearu-Levenshtein distance/similarity
* `Hamming()`: Hamming distance/similarity
* `OSA()`: Optimal String Alignment distance/similarity
* `LCS()`: Longest Common Subsequence distance/similarity

#### Token-based: 

Not yet implemented.

#### Hybrid token-character:

* `MongeElkan()`: Monge-Elkan measure

#### Other:

* `InVocabulary()`: Compares strings using a reference vocabulary. Useful for 
  comparing names.
* `Lookup()`: Retrieves distances/similarities from a lookup table

### Numeric measures:

* `Euclidean()`: Euclidean (L-2) distance
* `Manhattan()`: Manhattan (L-1) distance
* `Chebyshev()`: Chebyshev (L-∞) distance
* `Minkowski()`: Minkowski (L-p) distance

## Installation

You can install the latest release from [CRAN](https://CRAN.R-project.org) 
by entering:

``` r
install.packages("comparator")
```

The development version can be installed from GitHub using `devtools`:

``` r
# install.packages("devtools")
devtools::install_github("ngmarchant/comparator")
```

## Example

A measure can be instantiated by calling its constructor function. 
For instance, we can define a Levenshtein similarity measure that ignores 
differences in upper/lowercase characters as follows:

```{r lev}
measure <- Levenshtein(similarity = TRUE, ignore_case = TRUE)
```

A measure can be used to compare vectors element-wise as follows:

```{r elementwise}
x <- c("John Doe", "Jane Doe")
y <- c("jonathon doe", "jane doe")
elementwise(measure, x, y)

# shorthand for above
measure(x, y)
```

Pairwise comparisons are also supported using the following syntax:

```{r pairwise}
# compare each value in x with each value in y and return a similarity matrix
pairwise(measure, x, y)

# compare the values in x pairwise and return a similarity matrix
pairwise(measure, x, return_matrix = TRUE)
```
